# Getting and Cleaning Data : Peer Assessment 

## Purpose

You should create one R script called run_analysis.R that does the following.

1. Merges the training and the test sets to create one data set.
2. Extracts only the measurements on the mean and standard deviation for each measurement. 
3. Uses descriptive activity names to name the activities in the data set
4. Appropriately labels the data set with descriptive variable names. 
5. Creates a second, independent tidy data set with the average of each variable for each activity and each subject. 

1. Merges the training and the test sets to create one data set.
The source files are from: https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip 
After extraction, the folder UCI HAR Dataset contains train and test subfolders and other files.
The working directory is set to UCI HAR Dataset.

```{r}
setwd("H:/Rproga/UCI HAR Dataset")
# read initial data sets into temporary ones
# and check number of records read and columns
tempds1 <- read.table("train/X_train.txt") # X training set
tempds2 <- read.table("test/X_test.txt")  # X test set
nrow(tempds1); ncol(tempds1)
nrow(tempds2); ncol(tempds2)
# we  rbind() them into XDS and check the rows and cols
XDS <- rbind(tempds1, tempds2)
nrow(XDS); ncol(XDS)
# All ok, we reapeat for the labels data
tempds1 <- read.table("train/y_train.txt") # training set labels
tempds2 <- read.table("test/y_test.txt")  # test set labels
nrow(tempds1); ncol(tempds1)
nrow(tempds2); ncol(tempds2)
# we  rbind() them into LDS and check the rows and cols
LDS <- rbind(tempds1, tempds2)
nrow(LDS); ncol(LDS)
# All ok, we reapeat for the subjects data
tempds1 <- read.table("train/subject_train.txt") # training set labels
tempds2 <- read.table("test/subject_test.txt")  # test set labels
nrow(tempds1); ncol(tempds1)
nrow(tempds2); ncol(tempds2)
# we  rbind() them into SDS and check the rows and cols
SDS <- rbind(tempds1, tempds2)
nrow(SDS); ncol(SDS)
```

2. Extracts only the measurements on the mean and standard deviation for each measurement.

```{r}
setwd("H:/Rproga/UCI HAR Dataset")
# We read in the features table and look for varibale names with mean() and std()
# some varibales have Mean in them but we are sticking to what is required in 2. above
meanstd <- read.table("features.txt")
select_meanstd <- grep("mean()|std()", meanstd[,2])   # from second column
head(meanstd, 4)
nrow(meanstd)
str(select_meanstd)   # just checking
# We can now subset XDS 
XDS <- XDS[,select_meanstd]
Nmeanstd <- meanstd[select_meanstd,]   #subsetting meanstd
nrow(Nmeanstd);head(Nmeanstd,10)
# Check columns in XDS, then give corresponding names
ncol(XDS)
names(XDS) <- Nmeanstd[,2]
# Remove the () from names and convert to lower case
names(XDS) <-gsub("\\(|\\)", "", tolower(names(XDS)))
head(XDS,6)
```

3. Uses descriptive activity names to name the activities in the data set

```{r}
setwd("H:/Rproga/UCI HAR Dataset")
# We read in the activities table
ADS <- read.table("activity_labels.txt")
# we remove underscore and convert to lower case
ADS[,2] <- gsub("_", "", tolower((ADS[,2])))
ADS
# replace the numbers  by activity labels in LDS
nrow(LDS)
for (line in 1:nrow(LDS)){
  LDS[line,1] <- ADS[LDS[line,1],2]
}
nrow(LDS); head(LDS,55)
# Just give name to the column of LDS
names(LDS) <- "activity"
```

4. Appropriately labels the data set with descriptive variable names.

```{r}
# We name the column of subject data set SDS
names(SDS) <- "subject"
# We now join the three data sets
FinalDS <- cbind(SDS,LDS,XDS)
head(FinalDS,10)
# we save it
write.table(FinalDS, "H:/Rproga/UCI HAR Dataset/FirstTidy.txt", sep="\t")
```

5. Creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r}
TidyDS <- aggregate(FinalDS[,-c(1,2)], by = list(subject= FinalDS$subject,activity = FinalDS$activity), FUN= "mean")
nrow(TidyDS)
head(TidyDS, 10)
# we save it
write.table(TidyDS, "H:/Rproga/UCI HAR Dataset/SecondTidy.txt", row.name=FALSE, sep="\t")
```











                  





